---
title: "De la page à la carte - Université de Fribourg"
author: "Miguel Betti"
date: "29/04/2026"
output: html_document
---

#0. Préparation

```{r}
#I load the working data: Session -> Set working directory -> Choose directory... [my directory]
setwd("/Users/miguelbetti/Desktop/Git/DH_FRIBOURG")
```

```{r}
#I check that I am in the correct folder
getwd()
```

```{r}
#I install the necessary libraries and packages
if(!require("leaflet"))
install.packages("leaflet")
library(leaflet)
library(leaflet.extras)

if(!require("dplyr"))
install.packages("dplyr")
library(dplyr)

if(!require("htmlwidgets"))
install.packages
library(htmlwidgets)

if(!require("readxl"))
install.packages("readxl")
library(readxl)
```
#Data can be imported into R in a variety of formats, including CSV, JSON, plain text, and spreadsheet files. In this exercise, we will work with an Excel file (.xlsx), which we load using the read_excel() function from the readxl package. If you want to load your own data, simply change the path highlighted in green (e.g., "data/sevilla.xlsx").

```{r}
#Read the CSV files and create new objects
corpus <- read_excel("data/corpus.xlsx")
```

#Some places appear in more than one work. To avoid duplicates on the map, we group the data by location, so that each place is represented by a single row listing all the works that mention it:

```{r}
corpus_grouped <- corpus %>%
  group_by(place, lat, lng) %>%
  summarise(
    works = paste(title, collapse = "<br>"),
    .groups = "drop"
  )
```


#1. Create the map: "El Madrid de Lope de Vega" (or your own project)

```{r}
#Create the leaflet map object
corpus_map <- leaflet() %>%

  #1. Add a satellite imagery basemap from Esri
  addProviderTiles(providers$OpenStreetMap) %>%

  #2. Set the initial view: centered on Spain
  setView(lng = -3.7038, lat = 40.4150, zoom = 14) %>%

  #3. Circles
  addCircles(data = corpus_grouped,         
             lng = ~lng, lat = ~lat,
             group = "Works",
             color = "red", weight = 1,
             radius = 20,
             fillOpacity = 0.4,
             popup = ~paste0("<b>Place:</b> ", place, "<br>","<b>Work(s):</b><br>", works)) %>%

# Add a static legend
addControl(
  html = "<div style='background:white; padding:8px; border-radius:4px; border:1px solid #ccc;'>
            <b>Madrid de Lope de Vega (1562-1635)</b><br>
            <span style='color:red;'>●</span> Places mentioned
          </div>",
  position = "bottomright"
)

#Print the map to the RStudio Viewer pane
corpus_map
```



#Save the map as a standalone HTML file in your working directory
```{r}
saveWidget(corpus_map, file="corpus_map.html")
browseURL("corpus_map.html")
```

#Excercices

1. Load your own corpus, if you have one.

2. Adjust the zoom to focus on Sevilla or Spain. 

3. Modify the labels.

4. Add a toggle control.





