---
title: "De la page à la carte - Université de Fribourg"
author: "Miguel Betti"
date: "29/04/2026"
output: html_document
---

#O. Setup

```{r}
#Load the working data: Session -> Set working directory -> Choose directory... [my directory]
setwd("/Users/miguelbetti/Desktop/Git/DH_FRIBOURG")
```

```{r}
#I check that I am in the correct folder
getwd()
```

```{r}
#Install the necessary libraries and packages
if(!require("leaflet"))
install.packages("leaflet")
library(leaflet)
library(leaflet.extras)

if(!require("dplyr"))
install.packages("dplyr")
library(dplyr)

if(!require("htmlwidgets"))
install.packages
library(htmlwidgets)

```


#I. Trial: my first map with leaflet

```{r}
#I create my first trial map
map <- leaflet() %>% addTiles() %>% 
    #Set the position and the zoom on Fribourg
      setView(7.15, lat = 46.80, zoom = 13)
#Open the map
map
```

```{r}
#I add a marker with a popup: Université de Fribourg - Beauregard
map <- leaflet() %>% addTiles() %>% 
      setView(7.15, lat = 46.80, zoom = 13) %>% 
      addMarkers(lng = 7.146065736643453, lat = 46.802408523089895)
map
```

```{r}
#I add a a popup with more information
map <- leaflet() %>% addTiles() %>% 
      setView(7.15, lat = 46.80, zoom = 13) %>% 
      addMarkers(lng = 7.146065736643453,
                 lat = 46.802408523089895,
                 popup = "<b>Université de Fribourg</b><br>
                          Site Beauregard (BQC)<br>
                          Avenue de Beauregard 11<br>
                          CH-1700 Fribourg<br>
                          Salle 2.525")
map
```

#2. My first literary map: "El mapa del Quijote"

#Data can be imported into R in a variety of formats, including CSV, JSON, plain text, and spreadsheet files. In this exercise, we will work with two CSV files already prepared, containing all the places mentioned in the two parts of El Quijote by Miguel de Cervantes.

```{r}
##I create two objects with my data (books, toponyms and coordinates). I split them into two groups because I want to display them separately on the map.
parte_I <- read.csv("data/Parte_I.csv") #Primera parte del Quijote
parte_II <- read.csv("data/Parte_II.csv") #Segunda parte del Quijote
```


```{r}
#Create the leaflet map object
dq_map <- leaflet() %>%

  #1.Add a satellite imagery basemap from Esri (not the basic Leaflet one)
  addProviderTiles(providers$Esri.WorldImagery) %>%

  #2.Set the initial view: centered on the Mediterranean
setView(lng = 14, lat = 38, zoom = 4) %>%
  
  #3.Add markers

  #PART I markers
  addCircles(data = parte_I,          #Use the parte_I data frame as the source
             lng = ~lng, lat = ~lat,  #Map the lng and lat columns to coordinates
             group = "Primera parte (1605)",       #Layer group name for the toggle control
             radius = ~sqrt(menciones) * 10,       #Circle size scaled by mention count "menciones"
             color = "blue", weight = 5,           #color and thickness
             fill = TRUE, fillColor = "blue", fillOpacity = 0.5,  # Opacity
             popup = ~paste0("Lugar: ", lugar, "<br>Menciones: ", menciones)) %>%  #click popup with place name and count

  #PART II markers
  addCircles(data = parte_II,         
             lng = ~lng, lat = ~lat,
             group = "Segunda parte (1615)",
             radius = ~sqrt(menciones) * 1000,
             color = "red", weight = 5,
             fill = TRUE, fillColor = "red", fillOpacity = 0.5,
             popup = ~paste0("Lugar: ", lugar, "<br>Menciones: ", menciones)) %>%

  #Add a toggle control so users can show/hide each part's layer independently
  addLayersControl(
    baseGroups = c("Don Quijote de la Mancha"),              #the basemap label
    overlayGroups = c("Primera parte (1605)", "Segunda parte (1615)"),  #toggleable layers
    options = layersControlOptions(collapsed = TRUE))        #control starts collapsed

#Open the map
dq_map
```


```{r}
#I can save the map as a HTML file in my working directory
saveWidget(dq_map, file="Cervantes_map.html")
browseURL("Cervantes_map.html")
```

#Exercice

1. Change the basemap:
  # addProviderTiles(providers$OpenStreetMap) %>%         # Classic street map
  # addProviderTiles(providers$Esri.NatGeoWorldMap) %>%   # National Geographic
  # addProviderTiles(providers$CartoDB.Positron) %>%      # Minimal light grey

2. Zoom to Spain (coordinates? Zoom size?)

3. Play with colour and circle size

4. Add information to the pop_up (obra? tipo?)



